An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2022-12-30 article/article.tex
rerun script with figures

\begin{figure}[ht]
  \begin{center}
    \input{latex_files/images/chapter3/example_gt.tex}
    \input{latex_files/images/chapter3/example_post.tex}
  \end{center}
  \caption[Histogram of distances at the one-dimensional example.]{Histogram of
    distances and visualization of a specific region.}
  \label{fig:running_example_romc_inference}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ c|c|c|c|c }
\hline
& \(\mu_{\theta_1}\) & \(\sigma_{\theta_1}\) & \(\mu_{\theta_2}\) & \(\sigma_{\theta_2}\) \\
\hline \hline
Rejection ABC & 0.516 & 0.142 & 0.07 & 0.172 \\
\hline
ROMC (gradient-based) & 0.501 & 0.142 & 0.033 & 0.169 \\
\hline
ROMC (Bayesian optimization) & 0.494 & 0.16 & 0.086 & 0.167 \\
\hline
ROMC (Neural Network) & 0.491 & 0.138 & 0.04 & 0.172 \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of the samples obtained from the estimated
  posterior with (a) Rejection sampling and (b) the different versions
  of ROMC. We observe that the obtained samples share similar
  statistics along all methods. \label{tab:ma2}}
\end{table}

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1.png}
        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.png}
    \end{center}
  \caption[The acceptance region of a specific deterministic simulator.]{The acceptance region in a specific optimization problem. In the left figure the region obtained with gradient-based optimizer and in the right one with Bayesian Optimization.}
  \label{fig:ma2_5}
\end{figure}

\begin{figure}[ht]
  \begin{center}
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_rejection.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_romc.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.tex}
    }\\
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_rejection.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_romc.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex}
    }
    \resizebox{.24\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.tex}
    }
    \end{center}
    \caption[MA2 example, evaluation of the marginal
    distributions.]{Histogram of the marginal posterior distributions
      using three different inference approaches; (a) in the first
      row, the samples are obtained using Rejection ABC sampling (b)
      in the second row, using ROMC with a gradient-based optimizer
      and (c) in the third row, using ROMC with Bayesian optimization
      approach. The vertical (red) line represents the samples mean
      \(\mu\) and the horizontal (black) the standard deviation
      \(\sigma\).}
  \label{fig:ma2_3}
\end{figure}

\begin{figure}[ht]
  \begin{center}
    \resizebox{.32\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_romc_posterior.tex}
    }
    \resizebox{.32\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_romc_posterior_bo.tex}
    }
    \resizebox{.32\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/mae2_romc_posterior_nn.tex}
    }
    \end{center}
    \caption[MA2 example, posterior distribution.]{The unnormalized posterior distribution using the ROMC method with (a) a gradient-based optimization (b) Bayesian Optimization (c) gradient-based with a Neural Network as a surrogate model.}
  \label{fig:ma2_4}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \resizebox{.49\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/exec_time_solve.tex}
    }
    \resizebox{.49\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/exec_time_regions.tex}
    }
    \end{center}
    \caption[Execution time exploiting Parallelize]{Comparison
      between parallel and sequential execution of ROMC. We observe
      that the parallel version runs almost 5 times faster.}
      \label{fig:exec_parallel}
\end{figure}

>>>>

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{latex_files/images/chapter3/example_gt.pdf}
    \includegraphics[width=0.49\textwidth]{latex_files/images/chapter3/example_posterior.pdf}
    % \input{latex_files/images/chapter3/example_gt.tex}
    % \input{latex_files/images/chapter3/example_post.tex}
  \end{center}
  \caption[Histogram of distances at the one-dimensional example.]{Histogram of
    distances and visualization of a specific region.}
  \label{fig:running_example_romc_inference}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{ c|c|c|c|c }
\hline
& \(\mu_{\theta_1}\) & \(\sigma_{\theta_1}\) & \(\mu_{\theta_2}\) & \(\sigma_{\theta_2}\) \\
\hline \hline
Rejection ABC & 0.516 & 0.142 & 0.07 & 0.172 \\
\hline
ROMC (gradient-based) & 0.501 & 0.142 & 0.033 & 0.169 \\
\hline
ROMC (Bayesian optimization) & 0.501 & 0.17 & 0.087 & 0.175 \\
\hline
ROMC (Neural Network) & 0.491 & 0.138 & 0.04 & 0.172 \\
\hline
\end{tabular}
\end{center}
\caption{Comparison of the samples obtained from the estimated
  posterior with (a) Rejection sampling and (b) the different versions
  of ROMC. We observe that the obtained samples share similar
  statistics along all methods. \label{tab:ma2}}
\end{table}

\begin{figure}[ht]
    \begin{center}
        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1.pdf}
        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.pdf}
    \end{center}
  \caption[The acceptance region of a specific deterministic simulator.]{The acceptance region in a specific optimization problem. In the left figure the region obtained with gradient-based optimizer and in the right one with Bayesian Optimization.}
  \label{fig:ma2_5}
\end{figure}

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.pdf}\\
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.pdf}
    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.pdf}
    \end{center}
    \caption[MA2 example, evaluation of the marginal
    distributions.]{Histogram of the marginal posterior distributions
      using three different inference approaches; (a) in the first
      row, the samples are obtained using Rejection ABC sampling (b)
      in the second row, using ROMC with a gradient-based optimizer
      and (c) in the third row, using ROMC with Bayesian optimization
      approach. The vertical (red) line represents the samples mean
      \(\mu\) and the horizontal (black) the standard deviation
      \(\sigma\).}
  \label{fig:ma2_3}
\end{figure}

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior.pdf}
    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_bo.pdf}
    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_nn.pdf}
    \end{center}
    \caption[MA2 example, posterior distribution.]{The unnormalized posterior distribution using the ROMC method with (a) a gradient-based optimization (b) Bayesian Optimization (c) gradient-based with a Neural Network as a surrogate model.}
  \label{fig:ma2_4}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/exec_time_solve.pdf}
    \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/exec_time_regions.pdf}
    \end{center}
    \caption[Execution time exploiting Parallelize]{Comparison
      between parallel and sequential execution of ROMC. We observe
      that the parallel version runs almost 5 times faster.}
      \label{fig:exec_parallel}
\end{figure}

### 2022-12-30 article/latex_files/images/images/chapter4/mae2_hist_t2_romc_nn.tex
rerun script with figures

title={ROMC (BO) - \(\displaystyle \theta_2\)},

@@

\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.0617545870506399);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.494036696405118);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.308772935253199);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.864564218708957);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.11158256691152);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,0.802809631658317);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.49395771225155);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.71306067268783);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.17926096548025);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.06759941441516);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,1.5396445345721);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,1.91609586839403);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,1.79258669429275);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,1.59403798219366);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,1.58667684276459);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,1.52979906798699);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.15183142210051);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.06111444803853);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.47292129320938);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.85529643854941);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.86505006309548);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,2.67369166423286);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,1.96552066170352);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,1.29148734508544);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,1.68849292724457);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,1.00470598608042);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.703294190256297);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.703294190256294);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,1.2056471832965);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.200941197216084);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.100470598608042);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0);

@@

0.100012161271839 0
0.100012161271839 3

@@

\addlegendentry{$\mu = 0.100$}

@@

-0.108452084584719 1
0.278478839382765 1

@@

\addlegendentry{$\sigma = 0.176$}

>>>>

title={ROMC (Neural Network) - \(\displaystyle \theta_2\)},

@@

\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0557676847180302);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.134391393873674);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.232670307676795);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.421300598168856);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.599565962284508);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.819487590375745);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.02995995908051);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.25665578855782);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.57888096147667);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.58957701821668);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.49600073368982);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.52507821911024);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.62625669151916);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,1.92752590520997);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.09127677525018);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.20245981684059);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.32191116112789);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.50351769805879);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.54297681090209);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.14701105787528);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.97050861768377);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.65952774547258);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.41333954655085);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.1362319294878);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,0.992545439733498);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.95508724072649);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.820509887123889);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.693510954529521);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.644349818384951);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.49941685245109);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.304905545384957);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.137853649275808);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.144181418842862);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.0965680200621962);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0871794625561498);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.087179462556149);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0871794625561498);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.107297800069107);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0603550125388724);

@@

0.0396457758194439 0
0.0396457758194439 3

@@

\addlegendentry{$\mu = 0.040$}

@@

-0.170712448191162 1
0.207933154993939 1

@@

\addlegendentry{$\sigma = 0.172$}
